---
title: "4 Data Importing and Tidy Data"
subtitle: "Personal Notes"
author: "Max Hachemeister"
date: last-modified
date-format: iso
format: gfm
toc: true
editor: source
---

## Prerequisites

```{r}
# The "setup" makes the code chunk execute before any other, when this file is accessed again.

#| label: setup
#| include: false


# packages don't get saved between sessions, so need to install them each time
## install.packages(c(
##   "dplyr",
##   "ggplot2",
##   "readr",
##   "tidyr",
##   "nycflights23",
##   "fivethirtyeight",
##   "markdown"
##   ))

# and then open the packages
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(nycflights23)
library(fivethirtyeight)
library(markdown)
```

## Introduction

#### !Clarity

> You will see that having data stored in “tidy” format is about more than just what the everyday definition of the term “tidy” might suggest: having your data “neatly organized.” Instead, we define the term “tidy” as it’s used by data scientists who use R, outlining a set of rules by which data is saved.

I'd like these sentences to be more homogenous, by using the same sentence deliminator for both. So either two colons, or two commas, or so.

```{r}
library(readr)
dem_score <- read_csv("https://moderndive.com/data/dem_score.csv")
View(dem_score)
```

## 4.2 Tidy Data

#### LC4.1

> What are common characteristics of “tidy” data frames?

-   Each Column is a Variable

    -   if columns are named like "x2_height" "x2_height" "x3_height", then it's better to make another column that refers to those Xs

-   Each Row is an Observation

    -   this is less obvious, and I can't think of an example at the moment. Gotta go fast!

-   Each field contains just one value

    -   yeah, these are the entries that have values like "33C/15F" or something.

#### LC4.2

> What makes “tidy” data frames useful for organizing data?

It's especially readable for machines, and therefore all of the tidyverse functions work best with tidy data.

### LC4.3

> Take a look at the `airline_safety` data frame included in the `fivethirtyeight` data package.
>
> This data frame is not in “tidy” format. How would you convert this data frame to be in “tidy” format, in particular so that it has a variable `fatalities_years` indicating the incident year and a variable `count` of the fatality counts?

```{r}
airline_safety_smaller <- airline_safety |>
  select(airline, starts_with("fatalities"))
airline_safety_smaller
```

Okay, we can see, that the columns have the naming xx_yy, so i can deduce that this needs to be split up.

So I want to pivot longer, an put the names to `fatalities_years` and the values to `count`. An I need to do this only on the columns with `fatalities` in the name.

```{r}
airline_safety_smaller |>
  pivot_longer(
    names_to = "fatalities_years",
    values_to = "count",
    cols = starts_with("fatalities")
  )
```

Now the values for `fatalities_years` are a bit ugly still, but that's not the task right now.

## 4.3 Case study: democracy in Guatemala

#### !clarity

I think introducing the `names_transform` argument, which only works with the `list` function is confusing. Yeah it's the elegant way, but this can also be achieved with tools we've already learned and `mutate` is also just one line of code. Ah yeah, doing this with mutate is also a new concept, we haven't really talked about converting column types very much.

#### LC4.4

> Convert the `dem_score` data frame into a “tidy” data frame and assign the name of `dem_score_tidy` to the resulting long-formatted data frame.

```{r}
dem_score |>
  pivot_longer(
    names_to = "year",
    values_to = "democracy_score",
    cols = -country,
    names_transform = list(year = as.integer)
  )
```

#### LC4.5

> Read in the life expectancy data stored at <https://moderndive.com/data/le_mess.csv> and convert it to a “tidy” data frame.

```{r}
# import data
le_mess <-
read_csv("https://moderndive.com/data/le_mess.csv")

# convert to tidy
le_mess_tidy <-
le_mess |>
  pivot_longer(
    names_to = "year",
    values_to = "life_expectancy",
    cols = -country,
    names_transform = list(year = as.integer)
  )
le_mess_tidy
```


