---
title: "5 Simple Linear Regression"
subtitle: "Personal Notes"
author: "Max Hachemeister"
date: last-modified
date-format: iso
format: gfm
toc: true
editor: source
---

## Prerequisites

### [Link to Chapter](https://moderndive.com/v2/regression.html)

```{r}
# The "setup" makes the code chunk execute before any other, when this file is accessed again.

#| label: setup
#| include: false

# packages don't get saved between sessions, so need to install them each time
## install.packages(c(
##   "tidyverse",
##   "moderndive",
##   "markdown"
##   ))

# and then open the packages
library(tidyverse)
library(moderndive)
library(markdown)
```

## 5.1 One numerical explanatory Variable

#### !clarity

Please explain why we also take `obesity_rate_2016` in the filtered dataframe?

### 5.1.1 Exploratory data analysis

Get raw data and select columns of interest

```{r}
UN_data_ch5 <-
  un_member_states_2024 |>
  select(
    iso,
    life_exp = life_expectancy_2022,
    fert_rate = fertility_rate_2022,
    obes_rate = obesity_rate_2016
  ) |>
  na.omit()
```

Take a random sample to get an intuition about the data.

```{r}
UN_data_ch5 |>
  slice_sample(n = 5)
```

Get a statistical summary of the data.

Here we just take the two main variables, as `iso` being type character isn't sensibly summarized by standard numeric functions and `obes_rate` is not of interest at the moment.

```{r}
UN_data_ch5 |>
  # I'm using the numbers of the columns here, because I'm to lazy to type them out
  select(2, 3) |>
  tidy_summary()

## this should also work directly inside the tidy_summary() function
UN_data_ch5 |>
  tidy_summary(columns = c(2,3))
```

Get the correlation (coefficient) for the two variables.

```{r}
# this can either be done via the cor() function within summarize()
UN_data_ch5 |>
  summarize(correlation = cor(fert_rate, life_exp))

# aswell as with the get_correlation() function from moderndive
UN_data_ch5 |>
  get_correlation(fert_rate ~ life_exp)
```

Make a scatter plot for the two variables.

```{r}
UN_data_ch5 |>
  ggplot(aes(life_exp, fert_rate)) +
  geom_point(alpha = .1) +
  labs(x = "Life Expectancy", y = "Fertility Rate")
```

Add a `geom_smooth()` to visually represent the correlation.

```{r}
UN_data_ch5 |>
  ggplot(aes(life_exp, fert_rate)) +
  geom_point(alpha = .1) +
  # here is the line
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Life Expectancy",
       y = "Fertility Rate",
       title = "Relationship between Life Expectancy and Fertility Rate")
```

#### LC5.1

> Conduct a new exploratory data analysis with the same outcome variable y being `fert_rate` but with `obes_rate` as the new explanatory variable x:
>
> This involves three things
>
> 1.  Looking at the raw data values.
>
> 2.  Computing summary statistics.
>
> 3.  Creating data visualizations.
>
> What can you say about the relationship between obesity rate and fertility rate based on this exploration?

Look at the raw data values:

```{r}
# check out the data
UN_data_ch5 |>
  view()

# take a random sample
UN_data_ch5 |>
  slice_sample(n = 5)
```

Compute summary statistics:

```{r}
# univariate statistical summary
UN_data_ch5 |>
  tidy_summary(columns = c(fert_rate, obes_rate))

# bivariate statistic, correlation coefficient
## version 1
UN_data_ch5 |>
  summarize(correlation_coefficient = cor(fert_rate, obes_rate))

## version 2
UN_data_ch5 |>
  get_correlation(formula = fert_rate ~ obes_rate)
```

The middle 50% of the countries have a fertility rate between 1.6 and 3.2 and an obesity rate between 9.6 and 20.6. The median fertility rate is 2.0, while that of the obesity rate is 20.6.

The fertility rate seems to correlate slightly negative with obesity rate.

Visualize the data:

```{r}
# make a scatterplot and fit a linear regression line
UN_data_ch5 |>
  ggplot(aes(x = obes_rate, y = fert_rate)) +
  geom_point(alpha = .1) +
  geom_smooth(method = "lm",
              formula = "y ~ x",
              se = FALSE) +
  labs(x = "Obesity Rate",
       y = "Fertiliy Rate",
       title = "The Relationship between Fertility Rate and Obesity Rate"
  )
```

Fertility Rate and Obesity Rate show a rather random distribution, with a main cluster around a fertility rate of 1.5 and a obesity rate of 25, while for fertility rates between 2 to 4 the obesity rate seems most random.

#### LC5.2

> What is the main purpose of performing an exploratory data analysis (EDA) before fitting a regression model?

-   A. To predict future values.

-   *B. To understand the relationship between variables and detect potential issues.*

-   C. To create more variables.

-   D. To generate random samples.

LC5.3

> Which of the following is correct about the correlation coefficient?

-   A. It ranges from -2 to 2.

-   B. It only measures the strength of non-linear relationships.

-   *C. It ranges from -1 to 1 and measures the strength of linear relationships.*

-   D. It is always zero.

### 5.1.2 Simple linear regression

Fit a model and get the regression coefficient:

```{r}
# fit the model and save it to an object
demographics_model <-
  lm(fert_rate ~ life_exp, data = UN_data_ch5)

# get the coefficient
coef(demographics_model)
```

#### LC5.4

> Fit a simple linear regression using `lm(fert_rate ~ obes_rate, data = UN_data_ch5)` where `obes_rate` is the new explanatory variable $x$. Learn about the “best-fitting” line from the regression coefficients by applying the [`coef()`](https://rdrr.io/r/stats/coef.html) function. How do the regression results match up with your earlier exploratory data analysis?

Fit a linear model and get the regression coefficients:

```{r}
# fit the linear model
lm_fert_obes <-
  lm(fert_rate ~ obes_rate, data = UN_data_ch5)

# get the coefficient
coef(lm_fert_obes)
```

The regression coefficient and the correlation coefficient are both negative, and also only represent a weak relationship between fertility rate and obesity rate.

#### LC5.5

> What does the intercept term $b_0$ represent in simple linear regression?

-   A. The change in the outcome for a one-unit change in the explanatory variable.

-   *B. The predicted value of the outcome when the explanatory variable is zero.*

-   C. The standard error of the regression.

-   D. The correlation between the outcome and explanatory variables.

#### LC5.6

> What best describes the “slope” of a simple linear regression line?

-   A. The increase in the explanatory variable for a one-unit increase in the outcome.

-   B. The average of the explanatory variable.

-   *C. The change in the outcome for a one-unit increase in the explanatory variable*.

-   D. The minimum value of the outcome variable.

#### LC5.7

> What does a negative slope in a simple linear regression indicate?

-   *A. The outcome variable decreases as the explanatory variable increases.*

-   B. The explanatory variable remains constant as the outcome variable increases.

-   C. The correlation coefficient is zero.

-   D. The outcome variable increases as the explanatory variable increases.

### 5.1.3 Observed/fitted values and residuals

#### !typo

In the mathematical notation "life_exp" has an extra "\\"

This should be the solution $\text{life}\_\text{exp}$

Also "\* arrow: \[...\]" seems not formatted

#### !clarity

The Value in the table is not the same as in the text. Probably due to default rounding. I'd prefer both to be the same

Get the fitted values and the residuals for all observations:

```{r}
get_regression_points(demographics_model)
```

#### LC5.8

> What is a “wrapper function” in the context of statistical modeling in R?

-   A. A function that directly fits a regression model without using any other functions.

-   *B. A function that combines other functions to simplify complex operations and provide a user-friendly interface.*

-   C. A function that removes missing values from a dataset before analysis.

-   D. A function that only handles categorical data in regression models.

#### LC5.9

> Generate a data frame of the residuals of the *Learning check* model where you used `obes_rate` as the explanatory x variable.

Generate data frame with regression points for `lm_fert_obes` model:

```{r}
get_regression_points(lm_fert_obes)
```

LC5.10

> Which of the following statements is true about the regression line in a simple linear regression model?

-   A. The regression line represents the average of the outcome variable.

-   *B. The regression line minimizes the sum of squared differences between the observed and predicted values.*

-   C. The regression line always has a slope of zero.

-   D. The regression line is only useful when there is no correlation between variables.

## 5.2 One categorical explanatory variable

#### !clarity

I'm having a hard time understanding this sentence:

> We use this data for basic regression again, but now using an explanatory variable x
>
> that is categorical, as opposed to the numerical explanatory variable model we used in the previous Section [5.1](https://moderndive.com/v2/regression.html#model1):
>
> 1.  A numerical outcome variable y (a country’s life expectancy) and
>
> 2.  A single categorical explanatory variable x (the continent that the country is a part of).

### 5.2.1 Exploratory data analysis

Create a data frame with columns of interest and without incomplete observations:

```{r}
gapminder2022 <-
  un_member_states_2024 |>
  select(country,
         life_exp = life_expectancy_2022,
         continent,
         gdp_per_capita) |>
  na.omit()
```

Get a first overview of the data:

```{r}
glimpse(gapminder2022)

gapminder2022 |>
  slice_sample(n = 5)
```

#### !clarity

here another function for slicing samples is used `sample_n()`, which has bee superseded by `slice_sample()`

Get some univariate summary statistics:

```{r}
gapminder2022 |>
  tidy_summary(columns = c(life_exp, continent))
```

#### !clarity

the values for median don't match. In the text it's 75.14, in the table it's 75.2 and according to my code it's 75.18.

Same for the mean.

Visualize the the distribution of `life_exp`:

```{r}
gapminder2022 |>
  ggplot(aes(life_exp)) +
  geom_histogram(binwidth = 5, color = "grey76") +
  labs(title = "Histrogram of distribution of worldwide life expectancies",
       x = "Life expectancy",
       y = "# Countries")
```

Show Histogram for each continent.

```{r}
gapminder2022 |>
  ggplot(aes(life_exp)) +
  geom_histogram(binwidth = 5, color = "grey76") +
  # change title accordingly
  labs(title = "Histrogram of distribution of life expectancies per continent",
       x = "Life expectancy",
       y = "# Countries") +
  # make a Histogram for each continent
  facet_wrap(~ continent, nrow = 2)
```

Make a side-by-side boxplot per continent:

```{r}
gapminder2022 |>
  ggplot(aes(continent, life_exp)) +
  geom_boxplot() +
  labs(title = "Distribution life expectancies by continent",
       # I think the x-axis label would be redundant, as I've mentioned
       # it in the title
       x = NULL,
       y = "Life expectancy")
```

Now get the mean and median life expectancy for each per continent:

```{r}
life_exp_by_continent <-
gapminder2022 |>
  group_by(continent) |>
  summarize(median = round(median(life_exp), 1),
            mean = round(mean(life_exp), 1))

life_exp_by_continent
```

#### !clarity

there seems to be a rounding error for the mean of Asia, as it sais 50.0 in the table of the book and 74.9 in my code.

Also the rounding function is not given in the code of the book.

Get the difference of mean life expectancy for each continent to the lowest mean life expectancy:

```{r}
life_exp_by_continent |>
  select(continent,
         mean) |>
  mutate("Difference versus Africa" = mean - min(mean))
```

#### LC5.11

> Conduct a new exploratory data analysis with the same explanatory variable x being `continent` but with `gdp_per_capita` as the new outcome variable y
>
> What can you say about the differences in GDP per capita between continents based on this exploration?

Steps of EDA:

-   select data/columns of interest

<!-- -->

-   look at the raw data

-   get univariate statistics / five-point summary

    -   get bivariate statistics / correlation coefficient

-   visualize distributions

Select data/columns of interest and look at the raw data:

```{r}
gdp_worldwide <-
gapminder2022 |>
  select(-life_exp)

gdp_worldwide |>
  glimpse()

gdp_worldwide |>
  slice_sample(n = 5)

```

Get univariate statistics:

```{r}
gdp_worldwide |>
  tidy_summary(columns = c(gdp_per_capita, continent))
```

The global median GDP per capita is around 6741.3, while the mean with circa 18472.8 lies rather upward, indicating a skew in that direction. The relatively high standard deviation implies a high variation, which the high range of the minimum and maximum of 259 and 240862 respectively reflect.

Visualize Distributions

Make a Histogram

```{r}
gdp_worldwide |>
ggplot(aes(gdp_per_capita)) +
  geom_histogram(binwidth = 1e4, color = "grey76") +
  labs(title = "Histogram of distribution of wordwide GPDs per capita",
       x = "GDP per capita",
       y = "# Countries")
```

With the distribution being very wide, the countries with GDP per capita below the median are not well represented in this histogram.

Make a histogram for each continent:

```{r}
gdp_worldwide |>
ggplot(aes(gdp_per_capita)) +
  geom_histogram(binwidth = 1e4, color = "grey76") +
  labs(title = "Histogram of distribution of GPDs per capita per continent",
       x = "GDP per capita",
       y = "# Countries") +
  facet_wrap(~continent)
```

The distribution of the lower GDP per capita is still unclear due to the outliers in Europe.

Make a side-by-side barplot:

```{r}
gdp_worldwide |>
  ggplot(aes(continent, gdp_per_capita)) +
  geom_boxplot() +
  labs(title = "Distribution of GDP per capita",
       x = "Continent",
       y = "GDP per capita")
```

Okay, with such a wide distribution it's hard to get the different groups into perspective. Europe and Asia seem have the most variation of GDP per capita, while Africa appears to have the least in relation.

Let's also compute the mean for each continent an add the column with the relative difference to the lowest mean.

```{r}
gdp_worldwide_diff_mean <-
  gdp_worldwide |>
    group_by(continent) |>
    summarize(median = median(gdp_per_capita),
              mean = mean(gdp_per_capita)) |>
    mutate("Difference of mean to minimal mean" = mean - min(mean))

gdp_worldwide_diff_mean
```

#### LC5.12

> When using a categorical explanatory variable in regression, what does the baseline group represent?

-   A. The group with the highest mean

-   *B. The group chosen for comparison with all other groups*

-   C. The group with the most data points

-   D. The group with the lowest standard deviation

#### !clarity

I'm confused about "variable in regression", when so far we've just been computing singular datapoints and not yet a linear regression. Is computing the difference of each mean to the lowest mean already part of linear regression?

### 5.2.2 Linear regression

Fit a linear regression model and get the coefficients

```{r}
lm_life_continent <-
  lm(life_exp ~ continent, data = gapminder2022)

coef(lm_life_continent)
```

#### !clarity

> See [Chapter 15](https://r4ds.had.co.nz/factors.html) of *R for Data Science* ([Grolemund and Wickham 2017](https://moderndive.com/v2/references.html#ref-rds2016)) for examples.

This is a reference to the deprecated first edition of that Book. Would it be sensible to update that to chapter 16 of the second edition?

#### !typo

> In a statistical modeling context, this is also known as a *dummy variable*. In our case, we consider the first such indicator variable 1Amer(x). This indicator function returns 1 if a country is in the Asia, 0 otherwise:

Amer does not refer to Asia. This is also like that in the corresponding mathjax expression.

#### LC5.13

> Fit a linear regression using `lm(gdp_per_capita ~ continent, data = gapminder2022)` where `gdp_per_capita` is the new outcome variable. Get information about the “best-fitting” line from the regression coefficients. How do the regression results match up with the results from your previous exploratory data analysis?

Fit a linear regression:

```{r}
lm_gdp_continent <-
  lm(gdp_per_capita ~ continent, data = gapminder2022)

coef(lm_gdp_continent)

```

This model represents the data as we've computed in LC5.11.

The Intercept is the mean GDP per capita of all african countries and hence all other coefficients are the difference in mean GDP per capita to that for each other continent, which can also be seen in `gdp_worldwide_diff_mean`.

```{r}
gdp_worldwide_diff_mean
```

#### LC5.14

> How many “offsets” or differences from the baseline will a regression model output for a categorical variable with 4 levels?

-   A. 1

-   B. 2

-   *C. 3*

-   D. 4

### 5.2.3 Observed/fitted values and residuals

Get the regression points from `lm_life_continent`:

```{r}
get_regression_points(lm_life_continent,
                      ID = "country")

# what happens without ID = "country" ?
get_regression_points(lm_life_continent)
```

Ah okay, because the model refers to a data frame, the `ID` argument can access other columns than those the model was fit with, so you can get the corresponding values from these, like `country` in this case.

#### !clarity

also here getting slightly other values and `Antigua and Barbaidos` seems a new value.

#### LC5.15

> Which interpretation is correct for a positive coefficient in a regression model with a categorical explanatory variable?

-   A. It indicates the baseline group.

-   B. It represents the mean value of the baseline group.

-   *C. The corresponding group has a higher response mean than the baseline’s.*

-   D. The corresponding group has a lower response mean than the baseline’s.

#### LC5.16

> Which of the following statements about residuals in regression is true?

-   *A. Residuals are the differences between the fitted and observed response values.*

-   B. Residuals are always positive.

-   C. Residuals are not important for model evaluation.

-   D. Residuals are the predicted values in the model.

#### LC5.17

> Using either the sorting functionality of RStudio’s spreadsheet viewer or using the data wrangling tools you learned in Chapter [3](https://moderndive.com/v2/wrangling.html#wrangling), identify the five countries with the five smallest (most negative) residuals? What do these negative residuals say about their life expectancy relative to their continents’ life expectancy?

Identify the countries with the five smallest (most negative) residuals:

```{r}
get_regression_points(lm_life_continent,
                      ID = "country") |>
  slice_min(residual, n = 5)
```

The negative residuals of these countries represents how much their life expectancy differs to the average of the other countries of the same continent.

#### LC5.18

> Repeat this process, but identify the five countries with the five largest (most positive) residuals. What do these positive residuals say about their life expectancy relative to their continents’ life expectancy?

Get the countries with the highest residuals in life expectancy:

```{r}
lm_life_continent |>
  get_regression_points(ID = "country") |>
  slice_max(residual, n = 5)
```

Positive residuals in this case represent that these countries have a higher life expectancy than the average of the other countries of the same continent.

## 5.3 Related topics

"associated" is a good term, to express a corellation instead of implying a causation

### 5.3.1 Correlation it not necessarily causation

#### !clarity

Isn't the HDI more a parameter or principal component than a measurable variable? And it also contains the life expectancy as an input variable. I don't think that's the definition of a confounding variable.

### 5.3.2 Best-fitting line

Compute the sum of least squares for the regression model of fertility rate explained by life expectancy1:

```{r}
demographics_model |>
  get_regression_points() |>
  mutate(residual_squared = residual^2) |>
  summarize(sum_squared_residuals = sum(residual_squared))

```

#### LC5.19

> Note in Figure [5.12](https://moderndive.com/v2/regression.html#fig:three-lines) there are 3 points marked with dots and:
>
> -   The “best” fitting solid regression line in blue
>
> -   An arbitrarily chosen dotted red line
>
> -   Another arbitrarily chosen dashed green line
>
> FIGURE 5.12: Regression line and two others.
>
> Compute the sum of squared residuals by hand for each line. Show that the regression line in blue has the smallest value of these three lines.

```{r}
# create tibble with values
line_points <-
  tribble(
    ~line, ~point, ~residual,
    "red", 1, -.5,
    "red", 2, -1.5,
    "red", 3, .5,
    "green", 1, 0,
    "green", 2, -.5,
    "green", 3, 2,
    "blue", 1, .5,
    "blue", 2, -1,
    "blue", 3, .5
  )

# calculate sum of squared residuals
line_points |>
  group_by(line) |>
  summarize(sum_squared_residuals = sum(residual^2))
```

Blue has the smallest sum of squared residuals with a value of $1.5$

### 5.3.3

Check out the `augment()` and `clean_names()` function.

```{r}
library(broom)
library(janitor)
demographics_model |>
  # augment adds the model calculations to the corresponding dataframe
  augment() |>
  # mutate_if rounds the values of all numeric columns to two digits
  mutate_if(is.numeric, round, digits = 2) |>
  # clean names takes away the prefix signs and converts to snake case
  clean_names()
```
