---
title:       "Hypothesis Testing"
author:      "Max Hachemeister"
date:        last-modified
date-format: iso
format:
  gfm:
    toc:     true
---

# Prerequisites

```{r}
#| label: setup
#| results: false

library(tidyverse)
library(moderndive)
library(infer)
library(nycflights23)
library(ggplot2movies)
theme_set(theme_light())
```

##### !error
*Many of the relevant concepts, ideas, and* we have already introduced many 
of the necessary concepts to understand hypothesis testing in Chapters 7 
and 8. 

##### !error
introduce the activity that motivates the simulation-based approach for 
two-sample problems, data collected from *Spotify[,] to* investigate whether
metal music is more popular than deep-house music. In Sections 9.3, 9.4, and
9.5 we

#### !error
9.1.1

part from the null hypothesis. We find this simplification convenient as we 
focus on the equal part of the null hypothesis only *and [it] becomes*
*clearer* that evidence against the null hypothesis may come only with 
values to the left of 3.6, hence


##### !error
The decision made by the authors to call a song “popular” if it is above 
the 75th percentile (3rd quartile) of popularity is arbitrary and could be 
changed to any other value.*)*

##### !clarity
So, based on our sample of nm=1000 metal tracks and *nf=1000* deep house 
tracks, the point estimate for pm−pd

is the difference in sample proportions

ˆpm−ˆ*pf*=0.563−0.529=0.034.

This difference in favor of metal songs of 0.034 (3.4 percentage points) is 

> Shouldn't these be $p_d$?

##### !clarity

the samples would be the nm=1000 metal songs and the *nf=1000* deep house songs.

> Shouldn't this be $n_d$?
>
> This is also a paragraph down below

In our case, how will the difference in sample proportions *ˆpm−ˆpf* vary due to sampling under H0? Recall from

```{r}
spotify_metal_deephouse <- 
  spotify_by_genre |> 
    filter(track_genre %in% c("metal", "deep-house")) |> 
    select(track_genre, artists, track_name, popularity, popular_or_not)
```

```{r}
obs_diff_prop <- 
  spotify_metal_deephouse |> 
    specify(formula = popular_or_not ~ track_genre, success = "popular") |>     calculate(stat = "diff in props", order = c("metal", "deep-house"))

null_distribution <- 
  spotify_metal_deephouse |> 
    specify(formula = popular_or_not ~ track_genre, success = "popular") |>     hypothesize(null = "independence") |> 
    generate(reps = 1000, type = "permute") |> 
    calculate(stat = "diff in props", order = c("metal", "deep-house"))
```


```{r}
null_distribution |> 
  visualize(bins = 25) +
  shade_p_value(obs_stat = obs_diff_prop, direction = "right")

null_distribution |> 
  get_p_value(obs_stat = obs_diff_prop, direction = "right")
  
```


```{r}
bootstrap_distribution <- 
  spotify_metal_deephouse |> 
    specify(formula = popular_or_not ~ track_genre, success = "popular") |>     generate(reps = 1000, type = "bootstrap") |> 
    calculate(stat = "diff in props", order = c("metal", "deep-house"))

percentile_ci <- 
  bootstrap_distribution |> 
  get_confidence_interval(level = 0.90, type = "percentile")
percentile_ci

bootstrap_distribution |> 
  visualize() +
  shade_ci(endpoints = percentile_ci)
  
```

##### !clarity
9.4.1

Going back to the *pennies* example from Chapter 8, say we wanted to test if the mean weight of all chocolate-covered almonds was equal to 3.5 grams or not. 

> I think this refers to the almonds.


##### !clarity
9.4.1

We would be testing the value of a “point” μ

, the mean weight in grams of all chocolate-covered almonds, as follows

*H0:μ=3.5vs HA:μ≠3.5*

> In the sections 9.1.1 the values were 3.6


#### LC9.1

> Why does the following code produce an error? In other words, what about
> the response and predictor variables makes this not a possible 
> computation with the `infer` package?

```{r}
#| error: true

library(moderndive)
library(infer)
null_distribution_mean <- 
  spotify_metal_deephouse |> 
  specify(formula = popular_or_not ~ track_genre, success = "popular") |> 
  hypothesize(null = "independence") |> 
  generate(reps = 1000, type = "permute") |> 
  calculate(stat = "diff in means", order = c("metal", "deep-house"))
```

This code gives an error because we are asking it to calculate means, but
the variables are binary, so there is no mean but only a proportion to
be calculated. Which is technically the mean of 1s and 0s, but it probably
makes a difference in the hypothesize step and such.

#### LC9.2 

> Why are we relatively confident that the distributions of the sample
> proportions will be good approximations of the population distributions
> of popularity proportions for the two genres?

We are relatively confident in our distribution of sample proportions
reflecting the population's distribution of sample proportions because our
sample is a random sample of that population and as we repeat the sampling
sufficiently often we will reach the point at which the mechanic of the central limit theorem comes into play; that is, if the sampling size is 
either large enough, or the sample is repeated often enought we will the resulting distribution of sample statistics will follow a normal
distribution approximating that of the population.

#### LC9.3

> Using the definition of $p$-value, write in words what the $p$-value
> represents for the hypothesis test comparing the popularity rates for
> metal and deep house.

The $p$-value in the given case represents how likely we would observe 
difference in popularity proportions from our sample if this difference
were assumed to be up to chance (random). The lower this value the more
likely it is that the difference in popularity is not just by chance.

#### LC9.4

> Describe in a paragraph how we used Allen Downey's diagram to coclude
> if a statistical difference existed between the popularity rate of metal
> and deep house for the Spotify example.

We basically followed the diagram by first calculating the observed 
effect/test statistic from our original sample. In the next step we 
defined the null hypotheses and simulated data that would arise under this
hypotheses. Then we related the observed value with those from the
simulated data so to observe how likely the observed value would be under
the null hypotheses, which is expressed through the $p$-value. If this
would seem very unlikely we could conclude that an alternative hypothesis
might be more plausible.

##### !error

9.5.3

However, this is not possible as there will always be the possibility of committing one of these *error[s]* when making a decision based on sample data.

#### LC9.5

> What is wrong about saying, "The defendant is innocent." based on the US
> system of criminal trials?

As the court is only trying to decide whether someone is guilty or not a
not guilty ruling just says that there are not enough evidence for the
guilt, but that does not mean there are any evidence for the innocence, 
which would be a different task to begin with.

#### LC9.6

> What is the purpose of hypothesis testing?

Hypothesis testing is foremost a structured way of furthering scientific
inquiry in a diligent and therefore productive manner by allowing only
those implications that seem probable beyond a reasonable doubt to shape
the understanding of things.

#### LC9.7

> What are some flaws with hypothesis testing? How could we alleviate
> them?

Hypothesis testing, even though structured and well thought out, is 
at the end a matter of study design and the rational choice of acceptable
uncertainty. I can be prone to either falsely rejecting a true null
hypothesis or accepting one that is actually false. Because we are mostly
sampling from much larger and diverse populations, our findings might not
be true even if the testing procedure itself was correctly conducted.
We could alleviate that with more robust sampling methods, larger samples,
or an better sampling design. Generally we should employ state of the art
tools an knowledge, but foremost be knowledgable about any tools we use for
inferring about the truth of our world.

#### LC9.8

> Consider two $\alpha$ significance levels of 0.1 and 0.01. Of the two,
> which would lead to a higher chance of commiting a Type I error?

A significance level ($\alpha$) of 0.1 would imply a higher chance of a 
Type I error, because this would translate to -- in colloquial terms -- 
a 10% chance of rejecting a actually true null hypothesis, while an
$\alpha$ of 0.01 would mean only a chance of 1% in that regard.

## 9.6 Case Study: Are Action or Romance Movies Rated Higher?

### Visualize
```{r}
# Boxplot
movies_sample |> 
  ggplot(aes(genre, rating)) +
  geom_boxplot() +
  labs(
    x = "Movie Genre",
    y = "IMDb Rating"
  )
  
# Faceted Histogram
movies_sample |>
  ggplot(aes(rating)) +
  geom_histogram(binwidth = .1) +
  facet_wrap(~genre)
```

### Summary statistics
```{r}
movies_sample |> 
  group_by(genre) |> 
  summarize(
    n = n(),
    mean_rating = mean(rating),
    std_dev = sd(rating)
  )
```

##### !clarity

> In the tables about the sampling scenario some symbols have \hat and
> some have \bar. Maybe homogenize that, if there are meant to mean the 
> same.

> EDIT
> Ah yeah, they are actually right, because $\bar x$ is the actual mean
> of a sample, and $\hat \mu$ is the estimator for the population based on
> $\bar x$

### `infer` workflow

```{r}
null_distribution_movies <-   
  movies_sample |> 
    specify(formula = rating ~ genre) |> 
    hypothesize(null = "independence") |> 
    generate(reps = 5000, type = "permute") |> 
    calculate(stat = "diff in means", order = c("Action", "Romance"))

obs_diff_mean_movies <- 
  movies_sample |> 
  specify(formula = rating ~ genre) |> 
  calculate(stat = "diff in means", order = c("Action", "Romance"))

null_distribution_movies |> 
  get_p_value(obs_stat = obs_diff_mean_movies, direction = "both")
  

null_distribution_movies |> 
  visualize() +
  shade_p_value(obs_stat = obs_diff_mean_movies, direction = "both")
```

#### LC9.9

> Conduct the same analysis comparing action movies versus romantic movies
> using the median rating instead of the mean rating. What was different
> and what was the same?

Okay, `infer` workflow lessgooo:
```{r}
null_distribution_median <- 
  movies_sample |> 
    specify(formula = rating ~ genre) |> 
    hypothesize(null = "independence") |> 
    generate(reps = 1000, type = "permute") |>
    calculate(stat = "diff in medians", order = c("Action", "Romance"))

obs_diff_median_movies <- 
  movies_sample |> 
  specify(formula = rating ~ genre) |> 
  calculate(stat = "diff in medians", order = c("Action", "Romance"))

null_distribution_median |> 
  get_p_value(obs_stat = obs_diff_median_movies, direction = "both")

null_distribution_median |> 
  visualize() +
  shade_p_value(obs_stat = obs_diff_median_movies, direction = "both")
```

The main difference were just two lines, in which I defined "median"
instead of "mean". However, the results were interestingly different;
with the null distribution being more dense in the center and the 
$p$-value even higher, so even less evidence for a difference in median
rating of romance and action movies.

#### LC9.10

> What conclusion can you make from viewing the faceted histogram looking
> at `rating` versus `genre` that you could not see when looking at the 
> boxplot?

Let's make the facet view (which I already did above):
```{r}
movies_sample |> 
  ggplot(aes(rating)) +
  geom_histogram() +
  facet_wrap(~ genre)
```
What I can mostly see that the box plots did not show is that the 
distribution of the sample values is not normal, so we could not
immediately conduct sensible comparisons between the two. But yeah, that
was true from the beginning I guess. That's what the $t$-distribution and
resampling are made for.

So we can see that the distribution of values is rather serrated, which
a box plot would not show.

#### LC9.11

> Describe in paragraph how we used Allen Downey's diagram to concule if
> a statistical difference existed between mean movie ratings for action
> an romance movies.

To follow Allen Downey's diagram, we first got our observed sample 
statistic from the sample. Then we simulated from that sample a population
in accordance with the null hypothesis, got the distribution of the test
statistic from this simulated data, and lastly compared the observed value
with this distribution to see how likely it could have arisen from the
randomness of the hypothesized null distribution.

#### LC9.12

> Why are we relatively confident that the distributions of the sample
> ratings will be good approximations of the population distributions of
> ratings for the two genres?

We are confident that our simulated sampling distributions are good 
approximations for the population distribution because, firstly, we are
resampling so many times that the central limit theorem comes into play,
which states that with enough sampling effort the sampling distribution of
a population parameter will reflect the mean and deviance of the true
population.

#### LC9.13

> Using the definition of $p$-value, write in words what the $p$-value
> represents for the hypothesis test comparing the mean rating of romance
> to action movies.

The $p$-value gives the probability of value as extreme, or 
more extreme as the observed one under the conditionins of the null 
hypothesis.

#### LC9.14

> What is the value of the $p$-value for the two-sided hypothesis test
> comparing the mean rating of romance to action movies?

In case of the online book it's 0.004

#### LC9.15

> Test your data-wrangling knowledge and EDA skills:
>
>   * Use `dplyr` and `tidyr` to create the necessary data frames focused
>     on only action and romance movies (but not both) from the `movies`
>     data frame in the `ggplot2movies` package.
>
>   * Make a boxplot and faceted histogram of this population data 
>     comparing  ratings of action and romance movies from IMDb.
>
>   * Discuss how these plots compare to the similar plots produced for 
>     the `movies_sample` data.

##### Create the data frame
```{r}
movies_action_romance <- 
  movies |> 
    #get only those movies that have either Action or Romance
    filter((Romance == 1 & Action == 0) | (Action == 1 & Romance == 0)) |> 
    #get only columns of interest
    select(title, year, rating, Action, Romance) |> 
    #put action and romance in one 'genre' column
    mutate(genre = case_when(Action == 1 ~ "action",
                             Romance == 1 ~ "romance"),
           # and drop the columns not needed anymore
           .keep = "unused")
    
movies_action_romance
```
##### Make box plots and faceted histograms
```{r}
# box plot
movies_action_romance |> 
  ggplot(aes(genre, rating)) +
  geom_boxplot()

# faceted histogram
movies_action_romance |> 
  ggplot(aes(rating)) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~ genre)
```

##### Describe the plot

These plots from the whole data aset appear quite evenly distributed, so
we could compare the the two samples directly to this extend. Though it
should be noted that the size of these samples are around 4000 for each
genre, or roughly 9000 in total. So this is not always feasible and we 
could approximate this with the resampling we did.

#### T-test
```{r}
movies_action_romance |> 
  t_test(formula = rating ~ genre,
         order = c("action", "romance"),
         alternative = "two-sided")

movies_sample |> 
  t_test(formula = rating ~ genre,
         order = c("Action", "Romance"),
         alternative = "two-sided")
```

##### !clarity

9.7.1 Theory-based approach

Based on the p-value = 0.005 we reject the null hypothesis, the average
rating for the Romance movies is not the same as the average rating for the
Action movies. This result is similar to the one calculated using the
simulation-based approach.

> In the sections before we set the significance level to 0.001. Which
> would mean we do not reject the null hypotheses.