---
title:       "Hypthesis Testing"
author:      "Max Hachemeister"
date:        last-modified
date-format: iso
format:
  gfm:
    toc:     true
---

# Prerequisites

```{r}
#| label: setup
#| results: false

library(tidyverse)
library(moderndive)
library(infer)
library(nycflights23)
library(ggplot2movies)
theme_set(theme_light())
```

##### !error
*Many of the relevant concepts, ideas, and* we have already introduced many 
of the necessary concepts to understand hypothesis testing in Chapters 7 
and 8. 

##### !error
introduce the activity that motivates the simulation-based approach for 
two-sample problems, data collected from *Spotify[,] to* investigate whether
metal music is more popular than deep-house music. In Sections 9.3, 9.4, and
9.5 we

#### !error
9.1.1

part from the null hypothesis. We find this simplification convenient as we 
focus on the equal part of the null hypothesis only *and [it] becomes*
*clearer* that evidence against the null hypothesis may come only with 
values to the left of 3.6, hence


##### !error
The decision made by the authors to call a song “popular” if it is above 
the 75th percentile (3rd quartile) of popularity is arbitrary and could be 
changed to any other value.*)*

##### !clarity
So, based on our sample of nm=1000 metal tracks and *nf=1000* deep house 
tracks, the point estimate for pm−pd

is the difference in sample proportions

ˆpm−ˆ*pf*=0.563−0.529=0.034.

This difference in favor of metal songs of 0.034 (3.4 percentage points) is 

> Shouldn't these be $p_d$?

##### !clarity

the samples would be the nm=1000 metal songs and the *nf=1000* deep house songs.

> Shouldn't this be $n_d$?
>
> This is also a paragraph down below

In our case, how will the difference in sample proportions *ˆpm−ˆpf* vary due to sampling under H0? Recall from

```{r}
spotify_metal_deephouse <- 
  spotify_by_genre |> 
    filter(track_genre %in% c("metal", "deep-house")) |> 
    select(track_genre, artists, track_name, popularity, popular_or_not)
```

```{r}
obs_diff_prop <- 
  spotify_metal_deephouse |> 
    specify(formula = popular_or_not ~ track_genre, success = "popular") |>     calculate(stat = "diff in props", order = c("metal", "deep-house"))

null_distribution <- 
  spotify_metal_deephouse |> 
    specify(formula = popular_or_not ~ track_genre, success = "popular") |>     hypothesize(null = "independence") |> 
    generate(reps = 1000, type = "permute") |> 
    calculate(stat = "diff in props", order = c("metal", "deep-house"))
```


```{r}
null_distribution |> 
  visualize(bins = 25) +
  shade_p_value(obs_stat = obs_diff_prop, direction = "right")

null_distribution |> 
  get_p_value(obs_stat = obs_diff_prop, direction = "right")
  
```


```{r}
bootstrap_distribution <- 
  spotify_metal_deephouse |> 
    specify(formula = popular_or_not ~ track_genre, success = "popular") |>     generate(reps = 1000, type = "bootstrap") |> 
    calculate(stat = "diff in props", order = c("metal", "deep-house"))

percentile_ci <- 
  bootstrap_distribution |> 
  get_confidence_interval(level = 0.90, type = "percentile")
percentile_ci

bootstrap_distribution |> 
  visualize() +
  shade_ci(endpoints = percentile_ci)
  
```

##### !clarity
9.4.1

Going back to the *pennies* example from Chapter 8, say we wanted to test if the mean weight of all chocolate-covered almonds was equal to 3.5 grams or not. 

> I think this refers to the almonds.


##### !clarity
9.4.1

We would be testing the value of a “point” μ

, the mean weight in grams of all chocolate-covered almonds, as follows

*H0:μ=3.5vs HA:μ≠3.5*

> In the sections 9.1.1 the values were 3.6


#### LC9.1

> Why does the following code produce an error? In other words, what about
> the response and predictor variables makes this not a possible 
> computation with the `infer` package?

```{r}
#| error: true

library(moderndive)
library(infer)
null_distribution_mean <- 
  spotify_metal_deephouse |> 
  specify(formula = popular_or_not ~ track_genre, success = "popular") |> 
  hypothesize(null = "independence") |> 
  generate(reps = 1000, type = "permute") |> 
  calculate(stat = "diff in means", order = c("metal", "deep-house"))
```

This code gives an error because we are asking it to calculate means, but
the variables are binary, so there is no mean but only a proportion to
be calculated. Which is technically the mean of 1s and 0s, but it probably
makes a difference in the hypothesize step and such.

#### LC9.2 

> Why are we relatively confident that the distributions of the sample
> proportions will be good approximations of the population distributions
> of popularity proportions for the two genres?

We are relatively confident in our distribution of sample proportions
reflecting the population's distribution of sample proportions because our
sample is a random sample of that population and as we repeat the sampling
sufficiently often we will reach the point at which the mechanic of the central limit theorem comes into play; that is, if the sampling size is 
either large enough, or the sample is repeated often enought we will the resulting distribution of sample statistics will follow a normal
distribution approximating that of the population.

#### LC9.3

> Using the definition of $p$-value, write in words what the $p$-value
> represents for the hypothesis test comparing the popularity rates for
> metal and deep house.

The $p$-value in the given case represents how likely we would observe 
difference in popularity proportions from our sample if this difference
were assumed to be up to chance (random). The lower this value the more
likely it is that the difference in popularity is not just by chance.

#### LC9.4

> Describe in a paragraph how we used Allen Downey's diagram to coclude
> if a statistical difference existed between the popularity rate of metal
> and deep house for the Spotify example.

We basically followed the diagram by first calculating the observed 
effect/test statistic from our original sample. In the next step we 
defined the null hypotheses and simulated data that would arise under this
hypotheses. Then we related the observed value with those from the
simulated data so to observe how likely the observed value would be under
the null hypotheses, which is expressed through the $p$-value. If this
would seem very unlikely we could conclude that an alternative hypothesis
might be more plausible.