---
title:       "Inference for Regression"
subtitle:    "My Personal Notes"
author:      "Max Hachemeister"
date:        last-modified
date-format: iso
format:
  gfm:
    toc:     true
---

# Prerequisites

```{r}
#| label: setup
#| results: false

library(tidyverse)
library(moderndive)
library(infer)
library(gridExtra)
library(GGally)
theme_set(theme_light())
```

## 10.1

### Tidy data
```{r}
UN_data_ch10 <- 
  un_member_states_2024 |> 
  select(country,
         life_exp = life_expectancy_2022,
         fert_rate = fertility_rate_2022) |> 
  na.omit()
UN_data_ch10
```

### Fit linear model
```{r}
linear_model <- 
  lm(fert_rate ~ life_exp, data = UN_data_ch10)
coef(linear_model)
```

### Plot data an model
```{r}
UN_data_ch10 |> 
  ggplot(aes(life_exp, fert_rate)) +
  geom_point() +
  geom_smooth(method = "lm",
              formula = y ~ x,
              se = FALSE,
              linewidth = 0.5) +
  labs(
    title = "Relationship of Fertility Rate and Life Expectancy",
    x = "Life Expectancy",
    y = "Fertility Rate",
  )
```

### Determine values for single observations
#### Get Row ID for France
```{r}
UN_data_ch10 |> 
  rowid_to_column() |> 
  filter(country == "France") |> 
  pull(rowid)
```

#### Get Fit Data for France
```{r}
linear_model |> 
  get_regression_points() |> 
  filter(ID == 57)
```

### Old Faithful Eruptions
#### Get Summary
```{r}
old_faithful_2024 |> 
  select(duration, waiting) |> 
  tidy_summary()
```

#### Visualize
```{r}
old_faithful_2024 |> 
  ggplot(aes(duration, waiting)) +
  geom_point(alpha = 0.5)
```

#### Fit Regression Model
```{r}
lm_faithful <- 
  lm(waiting ~ duration, data = old_faithful_2024)

coef(lm_faithful)
sigma(lm_faithful)
```

So we can see that for each minute of eruption duration the waiting time
increases by 0.371 minutes on average, while we expect the waiting time to
be off for about 20.37 minutes off of the calculated time, on average.

### Relating Basic Regression to Other Methods


#### Two-sample difference in means
Back to the `movies_sample`
```{r}
# Fit models and get infos
lm_movies <- 
  lm(rating ~ genre, data = movies_sample)

get_regression_table(lm_movies)
```

#### ANOVA
Now the `spotify_by_genre` example
```{r}
# Selects columns of interest and get a sample
spotify_for_anova <- 
  spotify_by_genre |> 
  filter(track_genre %in% c("country", "hip-hop", "rock")) |> 
  select(artists, track_name, popularity, track_genre)

spotify_for_anova |> 
  slice_sample(n = 5)
```

```{r}
# Check with Box plots
spotify_for_anova |> 
  ggplot(aes(track_genre, popularity)) +
  geom_boxplot()
```

```{r}
# Check the means for each
spotify_for_anova |> 
  group_by(track_genre) |> 
  summarize(mean_popularity = mean(popularity))
```

```{r}
# Fit linear model
lm_spotify <- 
  lm(popularity ~ track_genre, data = spotify_for_anova)

lm_spotify |> 
  get_regression_table()
```

```{r}
# Do an analysis of variance
aov(popularity ~ track_genre, data = spotify_for_anova) |> 
  anova()
```

#### LC10.1

> What does the error term $\epsilon$ in the linear model 
> $Y = \beta_0 + \beta_1 \cdot X + \epsilon$
> represent?

  * A. The exact value of the response variable.
  * B. The predicted value of the response variable based on the model.
  * *C. The part of the response variable not explained by the line.*
  * D. The slope of the linear relationship between $X$ and $Y$.
  
#### LC10.2

> Which of the following is a property of the least squares estimators
> $b_0$ and $b_1$?

  * A. They are biased estimators of the population parameters $b_0$ and
       $b_1$.
       
  * *B. They are linear combinations of the observed responses*
       *$y_1, y_2,\dots, y_n$.*
  
  * C. They are always equal to the population parameters $b_0$
       and $b_1$.
       
  * D. They depend on the specific values of the explanatory variable
       $X$ only.
       
#### LC10.3

> How can the difference in means between two groups be represented in a
> linear regression model?

  * A. By adding an interaction term between the groups and the response
       variable.
       
  * B. By fitting separate regression lines for each group and comparing
       their slopes.
  
  * *C. By including a dummy variable to represent the groups.*
  
  * D. By subtracting the mean of the group from the mean of the other and
       using this difference as the predictor.

## Theory-based Inference for Simple Linear Regression

Back to `old-faithful`


#### LC10.4

> In the context of a linear regression model, what does the null 
> hypothesis $H_0 : \beta_1 = 0$ represent?

  * *A. There is no linear association between the response and the*
       *explanatory variable.*
       
  * B. The difference between the observed and predicted values is zero.
  
  * C. The linear association between response and explanatory variable
       crosses the origin
       
  * D. The probability of committing a Type II Error is zero.
  
#### LC10.5

> Which of the following is an assumption of the linear regression model?

  * *A. The error terms $\epsilon_i$ are normally distributed with*
       *constant variance.*
       
  * B. The error terms $\epsilon_i$ have a non-zero mean.
  
  * C. The error terms $\epsilon_i$ are dependent on each other.
  
  * D. The explanatory variable must be normally distributed.
  
#### LC10.6

> What does it mean when we say that the slope estimator $b_1$ is a random
> variable?

  * A. $b_1$ will be the same for every sample taken from the population.
  
  * B. $b_1$ is a fixed value that does not change with different samples.
  
  * *C. $b_1$ can vary from sample to sample due to sampling variation.*
  
  * D. $b_1$ is always equal to the population slope $\beta_1$.
  
  
### There is a wrapper function for that

```{r}
lm_faithful |> 
  get_regression_table()

summary(lm_faithful)
```

##### !error

10.2.5

The statistic column contains the t-test statistic for b0 (first row)
and b1 (second row). If we focus on b1, the t

    -test statistic was constructed using the equation

t=b1−0SE(b1)=*11.594*

which corresponds to the hypotheses H0:β1=0
versus HA:β1≠0.

> The table gives 11.4 for the $t$-value of $b_1$.


##### !error

10.2.6

or all the observations *i=1,[dots],n.* Recall that the residual as defined
in Subsection 5.1.3, is the observed response minus the fitted value. If we
denote the residuals with the letter e we get:


## PU 10.2.6 Model fit and model assumptions