---
title:       "Inference for Regression"
subtitle:    "My Personal Notes"
author:      "Max Hachemeister"
date:        last-modified
date-format: iso
format:
  gfm:
    toc:     true
---

# Prerequisites

```{r}
#| label: setup
#| results: false

library(tidyverse)
library(moderndive)
library(infer)
library(gridExtra)
library(GGally)
theme_set(theme_light())
```

## 10.1

### Tidy data
```{r}
UN_data_ch10 <-
  un_member_states_2024 |>
  select(
    country,
    life_exp = life_expectancy_2022,
    fert_rate = fertility_rate_2022
  ) |>
  na.omit()
UN_data_ch10
```

### Fit linear model
```{r}
linear_model <-
  lm(fert_rate ~ life_exp, data = UN_data_ch10)
coef(linear_model)
```

### Plot data an model
```{r}
UN_data_ch10 |>
  ggplot(aes(life_exp, fert_rate)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, linewidth = 0.5) +
  labs(
    title = "Relationship of Fertility Rate and Life Expectancy",
    x = "Life Expectancy",
    y = "Fertility Rate",
  )
```

### Determine values for single observations
#### Get Row ID for France
```{r}
UN_data_ch10 |>
  rowid_to_column() |>
  filter(country == "France") |>
  pull(rowid)
```

#### Get Fit Data for France
```{r}
linear_model |>
  get_regression_points() |>
  filter(ID == 57)
```

### Old Faithful Eruptions
#### Get Summary
```{r}
old_faithful_2024 |>
  select(duration, waiting) |>
  tidy_summary()
```

#### Visualize
```{r}
old_faithful_2024 |>
  ggplot(aes(duration, waiting)) +
  geom_point(alpha = 0.5)
```

#### Fit Regression Model
```{r}
lm_faithful <-
  lm(waiting ~ duration, data = old_faithful_2024)

coef(lm_faithful)
sigma(lm_faithful)
```

So we can see that for each minute of eruption duration the waiting time
increases by 0.371 minutes on average, while we expect the waiting time to
be off for about 20.37 minutes off of the calculated time, on average.

### Relating Basic Regression to Other Methods


#### Two-sample difference in means
Back to the `movies_sample`
```{r}
# Fit models and get infos
lm_movies <-
  lm(rating ~ genre, data = movies_sample)

get_regression_table(lm_movies)
```

#### ANOVA
Now the `spotify_by_genre` example
```{r}
# Selects columns of interest and get a sample
spotify_for_anova <-
  spotify_by_genre |>
  filter(track_genre %in% c("country", "hip-hop", "rock")) |>
  select(artists, track_name, popularity, track_genre)

spotify_for_anova |>
  slice_sample(n = 5)
```

```{r}
# Check with Box plots
spotify_for_anova |>
  ggplot(aes(track_genre, popularity)) +
  geom_boxplot()
```

```{r}
# Check the means for each
spotify_for_anova |>
  group_by(track_genre) |>
  summarize(mean_popularity = mean(popularity))
```

```{r}
# Fit linear model
lm_spotify <-
  lm(popularity ~ track_genre, data = spotify_for_anova)

lm_spotify |>
  get_regression_table()
```

```{r}
# Do an analysis of variance
aov(popularity ~ track_genre, data = spotify_for_anova) |>
  anova()
```

#### LC10.1

> What does the error term $\epsilon$ in the linear model 
> $Y = \beta_0 + \beta_1 \cdot X + \epsilon$
> represent?

  * A. The exact value of the response variable.
  * B. The predicted value of the response variable based on the model.
  * *C. The part of the response variable not explained by the line.*
  * D. The slope of the linear relationship between $X$ and $Y$.
  
#### LC10.2

> Which of the following is a property of the least squares estimators
> $b_0$ and $b_1$?

  * A. They are biased estimators of the population parameters $b_0$ and
       $b_1$.
       
  * *B. They are linear combinations of the observed responses*
       *$y_1, y_2,\dots, y_n$.*
  
  * C. They are always equal to the population parameters $b_0$
       and $b_1$.
       
  * D. They depend on the specific values of the explanatory variable
       $X$ only.
       
#### LC10.3

> How can the difference in means between two groups be represented in a
> linear regression model?

  * A. By adding an interaction term between the groups and the response
       variable.
       
  * B. By fitting separate regression lines for each group and comparing
       their slopes.
  
  * *C. By including a dummy variable to represent the groups.*
  
  * D. By subtracting the mean of the group from the mean of the other and
       using this difference as the predictor.

## Theory-based Inference for Simple Linear Regression

Back to `old-faithful`


#### LC10.4

> In the context of a linear regression model, what does the null 
> hypothesis $H_0 : \beta_1 = 0$ represent?

  * *A. There is no linear association between the response and the*
       *explanatory variable.*
       
  * B. The difference between the observed and predicted values is zero.
  
  * C. The linear association between response and explanatory variable
       crosses the origin
       
  * D. The probability of committing a Type II Error is zero.
  
#### LC10.5

> Which of the following is an assumption of the linear regression model?

  * *A. The error terms $\epsilon_i$ are normally distributed with*
       *constant variance.*
       
  * B. The error terms $\epsilon_i$ have a non-zero mean.
  
  * C. The error terms $\epsilon_i$ are dependent on each other.
  
  * D. The explanatory variable must be normally distributed.
  
#### LC10.6

> What does it mean when we say that the slope estimator $b_1$ is a random
> variable?

  * A. $b_1$ will be the same for every sample taken from the population.
  
  * B. $b_1$ is a fixed value that does not change with different samples.
  
  * *C. $b_1$ can vary from sample to sample due to sampling variation.*
  
  * D. $b_1$ is always equal to the population slope $\beta_1$.
  
  
### There is a wrapper function for that

```{r}
lm_faithful |>
  get_regression_table()

summary(lm_faithful)
```

##### !error

10.2.5

The statistic column contains the t-test statistic for b0 (first row)
and b1 (second row). If we focus on b1, the t

    -test statistic was constructed using the equation

t=b1−0SE(b1)=*11.594*

which corresponds to the hypotheses H0:β1=0
versus HA:β1≠0.

> The table gives 11.4 for the $t$-value of $b_1$.


##### !error

10.2.6

or all the observations *i=1,[dots],n.* Recall that the residual as defined
in Subsection 5.1.3, is the observed response minus the fitted value. If we
denote the residuals with the letter e we get:


### Residual Diagnostics

#### Check Linearity
```{r}
# Get residuals from model.

fit_and_residuals <-
  get_regression_points(lm_faithful)

# Plot residuals over fitted values

fit_and_residuals |>
  ggplot(aes(waiting_hat, residual)) +
  geom_point() +
  geom_hline(yintercept = 0, col = "blue")
```

##### !error

Now, the observations in this dataset are only a subset of all the Old
Faithful geyser eruptions that happen during this time frame and most
or all of them are eruptions that *do[did] not happen[ed]* sequentially,
one after the next.

#### Check Normality
```{r}
# Plot residuals as histogram
fit_and_residuals |>
  ggplot(aes(residual, y = after_stat(density))) +
  geom_histogram() +
  geom_density()

# Plot residuals with qq plot
fit_and_residuals |>
  ggplot(aes(sample = residual)) +
  geom_qq() +
  geom_qq_line(color = "blue")
```

##### !clarity

```{r}
fit_and_residuals |>
  ggplot(aes(sample = residual)) +
  geom_qq() +
  geom_qq_line(color = "blue")
```

> Should have put `color = "blue"` in `geom_qq_line()` to match the plot.

#### Check Equality/Constancy

Also known as *homoskedasticity*.

```{r}
# Plot residuals against regressor variable

fit_and_residuals |>
  ggplot(aes(duration, residual)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0) +
  labs(
    x = "Duration",
    y = "Residual"
  )
```

#### LC10.7

> Use the `un_member_states_2024` data frame included in the `moderndive`
> package with response variable `fertility_rate_2022` and the regressor
> `hdi_2022`. Make sure to omit missing values.
>
>   * Use the `get_regression_points()` function to get the observed
>     observed values.
> 
>   * Perform a residual analysis and look for any systematic patterns in
>     the residuals. Ideally, there should be little to no pattern but
>     comment on what you find here.

##### Tidy Data
```{r}
# tidy data
un_tidy <-
  un_member_states_2024 |>
  na.omit() |>
  select(country, fert_rate = fertility_rate_2022, hdi = hdi_2022)

# Fit model
lm_un <-
  lm(fert_rate ~ hdi, data = un_tidy)

# Get regression points.
un_regpoints <-
  lm_un |>
  get_regression_points()
```

##### Residual Analysis
###### Linearity
```{r}
# Check linearity by plotting residuals over estimates

un_regpoints |>
  ggplot(aes(fert_rate_hat, residual)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "blue")

```

It looks almost non-linear but the downward slope is kind of corrected by
the increased spread of the points. So let's assume it's not blatantly
non-linear.

###### Independence

We could check the independence either with another term like time.
Or explain why we think that our sample surely is random.
So for the sake of the exercise I assume the sample to be random.

###### Normality

```{r}
# Plot histogram of residuals
un_regpoints |>
  ggplot(aes(residual, after_stat(density))) +
  geom_histogram() +
  geom_density(color = "blue")

# Plot qq-plot
un_regpoints |>
  ggplot(aes(sample = residual)) +
  geom_qq() +
  geom_qq_line(color = "blue")

```

Yeah both look close to what I think a normal distribution looks like.

##### Equality of variance / Homoskedasticity

```{r}
# Plot residuals against regressor variable
un_regpoints |>
  ggplot(aes(hdi, residual)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0)
```

The range of the residuals stay more or less within an value of 2 over the
hdi range, but at the higher hdi the points both aggregate but also some
more deviant values appear. Bou there seems no blatant heteroskedasticity.

##### Verdict

Overall the assumptions to properly fit a simple linear regression for
inference are met.

#### LC10.8

> In the context of linear regression, a `p_value` of near zero for the
> slope coefficient suggests which of the following?

  * A. The intercept is statistically significant at a 95% confidence
       level.
        
  * *B. There is strong evidence against the null hypothesis that the*
       *slope coefficient is zero, suggesting there exists a linear*
       *relationship between the explanatory and response variables.*
        
  * C. The variance of the response variable is significantly greater
       than the variance of the explanatory variable.
        
  * D. The residuals are normally distributed with mean zero and constant
       variance.
       
#### LC10.9

> Explain whether or not the residual plot helps assess each one of the
> following assumptions.

  * Linearity of the relationships between variables
  * Independence of the error terms
  * Normality of the error terms
  * Equality or constancy of variance

Visualizing the residual can help assess all the above assumption to some
extent. The only thing that might not easily be checked is the 
independence of the error terms, if the variable causing the dependence
has not been observed. However, if the sampling procedure can be deemed
sufficiently random the independence of the error terms can also be 
assumed.

  
#### LC10.10

> If the residual plot against fitted values shows a "U-shaped" pattern,
> what does this suggest?

  * A. The variance of the residuals is constant.
  * *B. The linearity assumption is violated.*
  * C. The independence assumption of violated.
  * D. The normality assumption is satisfied.
  
## 10.3 Simulation-based Inference for Simple Linear Regression

### `infer` workflow
#### Get sampling distribution estimand by bootstrapping. 
```{r}
bootstrap_slope <-
  old_faithful_2024 |>
  specify(formula = waiting ~ duration) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "slope")
bootstrap_slope
```


#### Check distribution
```{r}
bootstrap_slope |>
  visualize()
```

#### Get confidence interval from percentiles
```{r}
bootstrap_slope_ci <-
  bootstrap_slope |>
  get_confidence_interval(type = "percentile", level = 0.95)
bootstrap_slope_ci
```

#### Alternatively get confidence interval from standard error
```{r}
# Get observed value
observed_slope <-
  old_faithful_2024 |>
  specify(waiting ~ duration) |>
  calculate(stat = "slope")
observed_slope

# Calculate ci from that
se_ci <-
  bootstrap_slope |>
  get_confidence_interval(
    level = 0.95,
    type = "se",
    point_estimate = observed_slope
  )
se_ci
```

### Hypothesis test
#### Generate null distribution
```{r}
null_distribution <-
  old_faithful_2024 |>
  specify(waiting ~ duration) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute")
```

#### Get test statistic
```{r}
null_distribution_slope <-
  null_distribution |>
  calculate(stat = "slope")

null_distribution_slope |>
  visualize()
```

#### Get observed slope and compare with null distribution
```{r}
# Get observed slope
observed_slope <-
  old_faithful_2024 |>
  specify(waiting ~ duration) |>
  calculate(stat = "slope")

# Visualize null distribution and shade in the p-value of observed value
null_distribution_slope |>
  visualize() +
  shade_p_value(obs_stat = observed_slope, direction = "both")

# And get actual p-value
null_distribution_slope |>
  get_p_value(obs_stat = observed_slope, direction = "both")
```

#### LC10.11

> Repeat the inference but this time for the correlation coefficient
> instead of the slope. Note the implementation of `stat = "correlation"`
> in the `calculate()` function for the `infer` package.

##### `infer` workflow
###### Generate distribution and ci for correlation
```{r}
# Generate bootstrap and get correlation
bootstrap_corr <-
  old_faithful_2024 |>
  specify(waiting ~ duration) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "correlation")

# Get observed correlation
obs_corr <-
  old_faithful_2024 |>
  specify(waiting ~ duration) |>
  calculate(stat = "correlation")

# Visually check for normality
bootstrap_corr |>
  visualize()
## Looks normal

# Get confidence interval from distribution percentiles
bootstrap_corr_ci_perc <-
  bootstrap_corr |>
  get_confidence_interval(level = 0.95, type = "percentile")

# Get confidence interval from standard error
bootstrap_corr_ci_se <-
  bootstrap_corr |>
  get_confidence_interval(level = 0.95, type = "se", point_estimate = obs_corr)
# Visualize
bootstrap_corr |>
  visualize() +
  shade_confidence_interval(endpoints = bootstrap_corr_ci_perc) +
  shade_confidence_interval(endpoints = bootstrap_corr_ci_se, color = "khaki")

```

Both confidence intervals do not contain the value 0, we are therefore 95%
confident that there is a correlation between the duration and the time 
between eruptions, on average.

##### Hypothesis check
###### Generate null distribution and get observed value

```{r}
null_distribution_corr <-
  old_faithful_2024 |>
  specify(formula = waiting ~ duration) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "correlation")

obs_corr <-
  old_faithful_2024 |>
  specify(formula = waiting ~ duration) |>
  calculate(stat = "correlation")
```


###### Visualize and get p value
```{r}
null_distribution_corr |>
  visualize() +
  shade_p_value(obs_stat = obs_corr, direction = "both")

null_distribution_corr |>
  get_p_value(obs_stat = obs_corr, direction = "both")
```

We tested the null hypothesis that the duration of and time between 
eruptions are not correlated. The p-value of the observed correlation is 0
meaning that it is highly unlikely for the correlation to have been observed
due to sampling variation under the null hypothesis. We therefore reject 
the hypothesis of independence and therefore assume the time between 
and the duration of eruptions to be correlated.

#### LC10.12

> Why is it appropriate to use the bootstrap percentile method to 
> construct a 95% confidence interval for the population slope $\beta_1$
> in the Old Faithful data?

  * A. Because it assumes the slope follows a perfect normal 
       distribution.
       
  * B. Because it relies on resampling the residuals instead of the 
       original data.
       
  * C. Because it requires the original data to be uniformly distributed.
  
  * *D. Because it does not require the bootstrap distribution to be*
       *normally shaped.*
       
#### LC10.13

> What is the role of the permutation test in the hypothesis testing for
> the population slope $\beta_1$?

  * A. It generates new samples to confirm the confidence interval
       boundaries.
       
  * *B. It assesses whether the observed slope could have occurred by*
       *chance under the null hypothesis of no relationship.*
       
  * C. It adjusts the sample size to reduce sampling variability.
  
  * D. It ensures the residuals of the regression model are normally
       distributed.
  
#### LC10.14

> After generating a null distribution for the slope using `infer`, you
> find the $p$-value to be near 0. What does this indicate about the 
> relationship between `waiting` and `duration` in the Old Faithful data?

  * A. There is no evidence of a relationship between `waiting` and
       `duration`
       
  * B. The observed slope is likely due to random variation under the 
       null hypothesis.
       
  * *C. The observed slope is significantly different from zero, *
       *suggesting a meaningful relationship between `waiting` and *
       *`duration`*
  
  * D. The null hypothesis cannot be rejected because the $p$-value is
       too small.
       
## 10.4 The Multiple Linear Regression Model
### Data exploration
#### Tidy Data
```{r}
coffee_data <-
  coffee_quality |>
  select(
    aroma,
    flavor,
    moisture_percentage,
    continent_of_origin,
    total_cup_points
  ) |>
  mutate(continent_of_origin = factor(continent_of_origin))
coffee_data
```

#### Get Tidy Summary
```{r}
coffee_data |>
  tidy_summary()
```

#### Create a scatterplot matrix
```{r}
#| message: false

coffee_data |>
  GGally::ggpairs()
```

#### LC10.15

> In a multiple linear regression model, what does the coefficient
> $\beta_j$ represent?

  * A. The intercept of the model.
  
  * B. The standard error of the estimate.
  
  * C. The total variance explained by the model.
  
  * *D. The partial slope related to the regressor $X_j$, accounting*
       *for all other regressors.*
       
#### LC10.16

> Why is it necessary to convert `continent_of_origin` to a factor when
> preparing the `coffee_data` data frame for regression analysis?

  * A. To allow the regression model to interpret `continent_of_origin`
       as a numerical variable.
       
  * *B. To create dummy variables that represent different categories of*
       *`continent_of_origin`.*
       
  * C. To reduce the number of observations in the dataset.
  
  * D. To ensure the variable is included in the correlation matrix.
  
#### LC10.17

> What is the purpose of creating a scatterplot matrix in the context of
> multiple linear regression?

  * A. To identify outliers that need to be removed from the dataset.
  
  * B. To test for normality of the residuals.
  
  * *C. To examine linear relationships between all variable pairs and*
       *identify multicollinearity among regressors.*
       
  * D. To determine the appropriate number of dummy variables.
  
#### LC10.18

> In the multiple regression model for the `coffee_data`, what is the
> role of dummy variables for `continent_of_origin`?

  * A. They are used to predict the values of the numerical regressors.
  
  * *B. They modify the intercept based on the specific category of*
       *`continent_of_origin`.*
       
  * C. They serve to test the independence of residuals.
  
  * D. The indicate which observations should be excluded from the model.
  
## 10.5 Theory-based Inference for Multiple Linear Regression

##### !error
10.5.1

 If we change the set of regressors used in a model, the *least-square[s]* estimates and *its[their]* standard errors will likely change as well. 
 
> Or why is ot plural here when in the context before it was not?
 
### Fit and inspect diverse regression models.

#### Model 1
```{r}
lm_multi_coff_1 <-
  lm(
    total_cup_points ~ aroma + flavor + moisture_percentage,
    data = coffee_data
  )

coef(lm_multi_coff_1)
sigma(lm_multi_coff_1)
```

#### Model 2
```{r}
lm_multi_coff_2 <-
  lm(total_cup_points ~ aroma + moisture_percentage, data = coffee_data)

coef(lm_multi_coff_2)
sigma(lm_multi_coff_2)
```

```{r}
lm_multi_coff_1 |>
  get_regression_table()
```

##### !clarity

10.5.2
For example, the formula for a 95% confidence interval for β1 is given by *b1±q⋅SEb1(s)* where the critical value q is determined by the level of confidence required, the sample size used (n), and the corresponding
degrees.

> Is that the right expression? I thought this should be $b_1 \pm q \cdot SE(b_1)$. Or what does the $s$ do?

  * three partial slopes for aroma, flavor and moisture_content
    (b1,b2, and b3), and

  * three coefficients for the factor levels in the model 
    *(b02, b03, and b04)*.

> Ah okay, yeah the difference is $1$ and $01$, sure. Would it be wrong to
> name them $b4, b5, b6$?

### ANOVA Test
```{r}
anova(lm_multi_coff_2, lm_multi_coff_1)
```

```{r}
# fit models with all columns as predictors
lm_multi_coff_all <-
  lm(
    total_cup_points ~
      aroma + flavor + moisture_percentage + continent_of_origin,
    data = coffee_data
  )

anova(lm_multi_coff_1, lm_multi_coff_all)
```

### Model Diagnostics
```{r}
# Get fitted values from the final model.
fit_and_res_coff <-
  lm_multi_coff_all |>
  get_regression_points()

# Make diagnostic plots.
g1 <-
  fit_and_res_coff |>
  ggplot(aes(total_cup_points_hat, residual)) +
  geom_point() +
  geom_hline(yintercept = 0, col = "blue") +
  labs(
    x = "Fitted Values (Total Cup Points)",
    y = "Residual"
  )
g2 <-
  fit_and_res_coff |>
  ggplot(aes(sample = residual)) +
  geom_qq() +
  geom_qq_line(col = "blue", linewidth = .5)

library(patchwork)
g1 + g2
```

##### !error
10.5.5
(#fig:inference-for-regression-arrange, grid-arrange-plot-check)Residuals vs. fitted values plot and QQ-plot for the multiple regression model. 

> Caption of plots is broken, ah yeah and therefore the reference is also
> just '??'.

#### LC10.19

> Why is it essential to know that the estimators 
> ($b_0, b_1, \dots, b_p$) in multiple linear regression are unbiased?

  * A. It ensures that the variance of the estimators is always zero.
  
  * *B. It means that, on average, the estimators will equal the true*
       *population parameters they estimate.*
       
  * C. It suggests that the estimators have a standard error of zero.
       
  * D. It implies that the regression model will always have a perfect
       fit.
       
#### LC10.20

> Why do the least-squares estimates of coefficients change when different
> sets of regressors are used in multiple linear regression?

  * A. Because the coefficients are recalculated each time, irrespective
       of the regressors.
       
  * B. Because the residuals are always zero when regressors are changed.
  
  * *C. Because the value of each coefficient depends on the specific*
       *combination of regressors included in the model.*
       
  * D. Because all models with different regressors will produce identical
       estimates.
       
#### LC10.21

> How is a 95% confidence interval for a coefficient in multiple linear
> regression constructed?

  * *A. By using the point estimate, the critical value from the*
       *$t$-distribution, and the standard error of the coefficient.*
       
  * B. By taking the standard deviation of the coefficients only.
  
  * C. By resampling the data without replacement.
  
  * D. By calculating the mean of all coefficients.
  
#### LC10.22

> What does the ANOVA test for comparing two models in multiple linear
> regression evaluate?

  * A. Whether all regressors in both models have the same coefficients.
  
  * *B. Whether the reduced model is adequate of if the full model is*
       *needed.*
       
  * C. Whether the residuals of the two models follow a normal 
       distribution.
       
  * D. Whether the regression coefficients of one model are unbiased
       estimators.
       
## 10.6 Simulation-based Inference for Multiple Linear Regression

#### !error

> In this section heading "Inference" is spelled with and captial I, while
> in the section 10.5 it's spelled with a lowercase i.

### Getting the Observed Fitted Model

```{r}
observed_fit <-
  coffee_data |>
  specify(
    total_cup_points ~
      aroma + flavor + moisture_percentage + continent_of_origin
  ) |>
  fit()

observed_fit
```

### Get the Bootstrap Distribution

```{r}
bootstrap_lm_coffee_all <-
  coffee_data |>
  specify(
    total_cup_points ~
      aroma + flavor + moisture_percentage + continent_of_origin
  ) |>
  generate(reps = 1000, type = "bootstrap") |>
  fit()

bootstrap_lm_coffee_all |>
  visualize()
```

### Get Confidence Intervals

```{r}
confidence_intervals_lm_coffee_all <-
  bootstrap_lm_coffee_all |>
  get_confidence_interval(
    level = 0.95,
    type = "percentile",
    point_estimate = observed_fit
  )

bootstrap_lm_coffee_all |>
  visualize() +
  shade_ci(endpoints = confidence_intervals_lm_coffee_all)

```

### Hypothesis Testing

##### !error
> In the code example for the null distribution the dataset 
> `coffee_quality` is used, while in the text above the `coffe_data` is
> referenced.

#### Generate Null Distribution
```{r}
set.seed(22051989)

null_distribution_lm_all <-
  coffee_data |>
  specify(
    total_cup_points ~
      continent_of_origin + aroma + flavor + moisture_percentage
  ) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  fit()

null_distribution_lm_all
```

#### Check Observed Against Null Distribution
```{r}
null_distribution_lm_all |>
  visualize() +
  shade_p_value(obs_stat = observed_fit, direction = "two-sided")
```

#### Get P-Values From Observed in Null Distribution
```{r}
null_distribution_lm_all |>
  get_p_value(obs_stat = observed_fit, direction = "both")

```

#### LC10.23

> Why might one prefer to use simulation-based methods
> (e.g. bootstrapping) for inference in multiple linear regression?

  * A. Because simulation-based methods require larger sample sized than
       theory-based methods.
       
  * B. Because simulation-based methods are always faster to compute than
       theory-based methods.
       
  * C. Because simulation-based methods guarantee the correct model is
       used.
       
  * *D. Because simulation-based methods do not rely on the assumptions*
       *of normality or large sample sizes.*
       
#### LC10.24

> What is the purpose of constructing a boostrap distribution for the
> partial slopes in multiple linear regression?

  * A. To replace the original data with random numbers.
  
  * *B. To approximate the sampling distribution of the partial slopes*
       *by resampling with replacement.*
       
  * C. To calculate the exact values of the coefficients in the 
       population.
       
  * D. To test if the model assumptions are violated.
  
#### LC10.25

> If a 95% confidence interval for a partial slope in multiple linear
> regression includes 0, what does this suggest about the variable?

  * *A. The variable does not have a statistically significant*
       *relationship with the response variable.*
       
  * B. The variable is statistically significant.
  
  * C. The variable's coefficient estimate is always negative.
  
  * D. The variable was removed from the model during bootstrapping.
  
#### LC10.26

> In hypothesis testing for the partial slopes using permutation tests,
> what does it mean if an observed test statistic falls far to the right
> of the null distribution?

  * A. The variable is likely to have no effect on the response.
  
  * B. The null hypothesis should be accepted.
  
  * *C. The variable is likely statistically significant, and we should*
       *reject the null hypothesis.*
       
  * D. The observed data should be discarded.

  Test and some more of that stuff, that I might like in the midst of things.
  Ans some more line.